Linking hypothesis and number of response options modulate inferred scalar implicature rate

### 2018-10-31 manuscript.tex
added auxiliary analysis of 3-way interaction

Our primary goal in this study was to test whether the estimated
implicature rate in the experimental task is affected by the linking
hypothesis and the number of response options available to participants.
To this end, we only analyzed the critical trials (exhaustive and
scalar). In particular, we classified each data point from critical
trials as constituting an implicature (1) or not (0) under the Strong
and Weak link. Figure \ref{fig:implicatureRatePlot} shows the resulting
implicature rates by condition and link.

Visually, the Weak link tends to result in greater estimates of
implicature rates, especially in tasks with more response options. Under
the Strong link, this latter pattern is reversed: the binary and ternary
judgment tasks result in greater estimates of implicature rates than
with more response options.

@@

To analyze the effect of link and response options on inferred
implicature rate, we used a Bayesian binomial mixed effects model using
the R packge \enquote{brms} (Bürkner \& others, 2016) with uninformative
or weakly informative priors.\footnote{For more information about the
  default priors of the \enquote{brms} package, see
  \href{ftp://cran.r-project.org/pub/R/web/packages/brms/brms.pdf}{the
  brms package manual}.} The model predicted the log odds of implicature
over no implicature from fixed effects of \emph{response type} (binary,
ternary, quaternary, quinary -- dummy-coded with binary as reference
level), \emph{link} (strong vs.~weak -- dummy-coded with strong as
reference level), and trial type (exhaustive vs.~scalar -- dummy-coded,
with exhaustive as reference level), as well as their two-way and
three-way interactions. Following Barr, Levy, Scheepers, and Tily
(2013), we included the maximal random effects structure justified by
the design: random intercepts for items (cards) and participants, random
by-participant slopes for link, trial type, and their interaction, and
random by-item slopes for link, trial type, response type, and their
interactions. Since the number of response options was a between
participant variable we did not include random slopes of response
options for participants. Four chains converged after 2000 iterations
each (warmup = 1000). Table \ref{tab:modeltable} summarizes the mean
parameter estimates and their 95\% credible intervals. \(\hat{R}=1\) for
all estimated parameters. All the analytical decisions described here
were pre-registered\footnote{You can access our pre-registration at
  \url{https://aspredicted.org/tq3sz.pdf}}.

The model provided evidence for the following effects: First, there was
a main effect of trial type such that scalar trials resulted in greater
implicature rates than exhaustive trials (Mean Estimate = 6.09, 95\%
Credible Interval={[}1, 12.29{]}). Second, there was an interaction
between link and number of response options such that the quaternary
task (Mean Estimate = 14.03, 95\% Credible Interval={[}7.24, 21.88{]})
and the quinary task (Mean Estimate = 17.28, 95\% Credible
Interval={[}10.64, 25.80{]}) with a weak link resulted in greater
implicature rates \change{than with a strong link}. Finally, there was a three-way interaction between
link, trial type, and number of response options (Mean Estimate = -7.74,
95\% Credible Interval={[}-16.59, -0.16{]}). One interpretation of this
interaction is that the difference between the Weak and Strong link on
scalar trials in the quinary task was smaller than on exhaustive trials,
though we believe this is not too interesting, given that the binary
reference level implicature estimate was lower for exhaustive trials in
the first place. Crucially, both number of response options and link
affect the inferred implicature rate.

>>>>

Our primary goal in this study was to test whether the estimated
implicature rate in the experimental task is affected by the linking
hypothesis and the number of response options available to participants.
To this end, we only analyzed the critical trials (exhaustive and
scalar). In particular, we classified each data point from critical
trials as constituting an implicature (1) or not (0) under the Strong
and Weak link. Figure \ref{fig:implicatureRatePlot} shows the resulting
implicature rates by condition and link. \change{It is immediately obvious that there is variability in inferred implicature rate. In particular, the Weak link appears to result in greater estimates of
implicature rates in tasks with four or five response options, compared to the Strong link. For the binary and ternary task, the assumed link appears to play a much smaller role.}

@@

To analyze the effect of link and response options on inferred
implicature rate, we used a Bayesian binomial mixed effects model using
the R packge \enquote{brms} (Bürkner \& others, 2016) with weakly informative priors.\footnote{For more information about the
  default priors of the \enquote{brms} package, see
  \href{ftp://cran.r-project.org/pub/R/web/packages/brms/brms.pdf}{the
  brms package manual}.} The model predicted the log odds of implicature
over no implicature from fixed effects of \emph{response type} (binary,
ternary, quaternary, quinary -- dummy-coded with binary as reference
level), \emph{link} (strong vs.~weak -- dummy-coded with strong as
reference level), and trial type (exhaustive vs.~scalar -- dummy-coded,
with exhaustive as reference level), as well as their two-way and
three-way interactions. Following Barr, Levy, Scheepers, and Tily
(2013), we included the maximal random effects structure justified by
the design: random intercepts for items (cards) and participants, random
by-participant slopes for link, trial type, and their interaction, and
random by-item slopes for link, trial type, response type, and their
interactions. Since the number of response options was a between-participant variable we did not include random slopes of response
options for participants. Four chains converged after 2000 iterations
each (warmup = 1000). Table \ref{tab:modeltable} summarizes the mean
parameter estimates and their 95\% credible intervals. \(\hat{R}=1\) for
all estimated parameters. All the analytical decisions described here
were pre-registered\footnote{Our preregistration can be accessed at
  \url{https://aspredicted.org/tq3sz.pdf}}.

The model provided evidence for the following effects: First, there was
a main effect of trial type such that scalar trials resulted in greater
implicature rates than exhaustive trials (Mean Estimate = 6.09, 95\%
Credible Interval={[}1, 12.29{]}). Second, there was an interaction
between link and number of response options such that the quaternary
task (Mean Estimate = 14.03, 95\% Credible Interval={[}7.24, 21.88{]})
and the quinary task (Mean Estimate = 17.28, 95\% Credible
Interval={[}10.64, 25.80{]}) resulted in greater
implicature rates \change{with a weak link than with a strong link, but there was no evidence of a link-dependent difference in inferred implicature rate for the binary and ternary task}. Finally, there was a three-way interaction between
link, trial type, and number of response options, \change{driven by the binary/quinary contrast} (Mean Estimate = -7.74,
95\% Credible Interval={[}-16.59, -0.16{]}). \change{Simple effects analysis on only the binary and and quinary trials, separately for the exhaustive and scalar subset of the data, revealed that the three-way interaction is driven by a different effect of number of response options under the Weak vs Strong link for the two inference types. Specifically, on exhaustive trials, number of response options (2 vs.~5) only resulted in greater implicature rates under the Weak ($\beta$ = .2, $p <$ .0001), but not the Strong link ($\beta$ = -.8, $p <$ .82). In contrast, on scalar trials, number of response options (2 vs.~5) resulted in greater implicature rates under the Weak ($\beta$ = 3.6, $p <$ .005) link, and in lower implicature rates under the Strong link ($\beta$ = -4.0, $p <$ .0007).}

\change{In sum,} both number of response options and link
affected the inferred implicature rate\change{, as did the type of inference (exhaustive vs.~scalar)}.
